---
title: "austin_animal_shelter_adoption_analysis"
author: "Nathan Funckes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
require(devtools)
install_github("Displayr/flipTime")
library(flipTime)
library(stringr)
library(scales)
library(lubridate)
library(ggfittext)
```

## Read Data

```{r read data}

intakes <- read_csv(here::here("data","Austin_Animal_Center_Intakes.csv"))
outcomes <- read_csv(here::here("data","Austin_Animal_Center_Outcomes.csv"))
```


```{r rename variables in intial datasets}
outcomes_renamed <- outcomes%>%
  rename("id"="Animal ID","name"="Name","outcomeDateTime"="DateTime","outcomeMonthYear"="MonthYear","dateOfBirth"="Date of Birth","outcomeType"="Outcome Type","outcomeSubtype"="Outcome Subtype","animalType"="Animal Type","outcomeSex"="Sex upon Outcome","outcomeAge"="Age upon Outcome","breed"="Breed","color"="Color")

intakes_renamed <-intakes%>%
  rename("id"="Animal ID","name"="Name","intakeDateTime"="DateTime","intakeMonthYear"="MonthYear","foundLocation"="Found Location","intakeType"="Intake Type","intakeCondition"="Intake Condition","animalType"="Animal Type","intakeSex"="Sex upon Intake","intakeAge"="Age upon Intake","breed"="Breed","color"="Color")
```


```{r combine datasets only on entries in both}
combined <- merge(intakes_renamed,outcomes_renamed,by=c("id"="id","name"="name","animalType"="animalType","breed"="breed","color"="color"))
adoption_data <- combined
```



```{r}
adoption_data
```


transform intakeAge and outcomeAge into numerical variables
transform the dates into date variables

```{r transform intakeAge, outcomeAge, dateOfBirth into date variables}
adoption_data$intakeDateTime <- AsDateTime(adoption_data$intakeDateTime)
adoption_data$outcomeDateTime <- AsDateTime(adoption_data$outcomeDateTime)
adoption_data$dateOfBirth <- AsDate(adoption_data$dateOfBirth)
adoption_data
```


```{r split outcomeMonthYear and intakeMonthYear into separate Month Year variables}
adoption_data$intakeMonth <-month(AsDate(adoption_data$intakeMonthYear))
adoption_data$intakeYear <-year(AsDate(adoption_data$intakeMonthYear))
adoption_data$outcomeMonth <-month(AsDate(adoption_data$intakeMonthYear))
adoption_data$outcomeYear <-year(AsDate(adoption_data$outcomeMonthYear))
```


```{r get days between income and outcome}
adoption_data$timeBetweenIntakeAndOutcome <- (as.Date(adoption_data$outcomeDateTime) - as.Date(adoption_data$intakeDateTime))
```

```{r remove observation where the time between intake and outcome is less than 0 days}
adoption_data <- adoption_data%>%
  filter(timeBetweenIntakeAndOutcome > 0)
adoption_data
```


adoption_data%>%
  mutate(intakeAgeYears = function_name(intakeAge))
transform intakeAge and outcomeAge into numerical values
```{r}
# 
# 
# ageStringToMonths <- function(ageString){
#     if(grepl("month", ageString, fixed=TRUE)){
#       ageInt <- as.double(str_extract(ageString,"[0-9]*"),base=0L)
#     }else if(grepl("year", ageString, fixed=TRUE)){
#       ageInt <- as.double(str_extract(ageString,"[0-9]*"),base=0L)
#       ageInt <- ageInt*12
#     }else{
#       ageInt<- NULL
#     }
#   }
```



```{r create a varaiable for age at intake and outcome in months}
# adoption_data <- adoption_data %>%
#   mutate(
#     intakeAgeBins = case_when(
#       intakeAgeMonths == "months" ~ as.double(numb),
#       intakeAgeMonths == "years" ~ 12*as.double(numb),
#       intakeAgeMonths == "month" ~ as.double(numb),
#       intakeAgeMonths == "year" ~ 12*as.double(numb),
#       intakeAgeMonths == "days" ~ 0,
#       intakeAgeMonths == "day" ~ 0
#   ))
# 
# adoption_data <- adoption_data %>%
#   mutate(
#     outcomeAgeBins = case_when(
#       intakeAgeMonths == "months" ~ as.double(numb),
#       intakeAgeMonths == "years" ~ 12*as.double(numb),
#       intakeAgeMonths == "month" ~ as.double(numb),
#       intakeAgeMonths == "year" ~ 12*as.double(numb),
#       intakeAgeMonths == "days" ~ 0,
#       intakeAgeMonths == "day" ~ 0
#   ))

```



```{r remove dulpicate ids}
#ids <- hashmap(keys=unique(adoption_data$id),values=0.0)

#ids['A141142']
  
```

```{r create age bins}

adoption_data <- adoption_data %>%
  separate(intakeAge, c("numb", "type")) %>% 
  mutate(
    intakeAgeBins = case_when(
      type == "day" ~ "Less than 6 months",
      type == "days" ~ "Less than 6 months",
      type == "week" ~ "Less than 6 months",
      type == "weeks" ~ "Less than 6 months",
      type == "month" ~ "Less than 6 months",
      type == "months" & strtoi(numb)< 6 ~ "Less than 6 months",
      type == "months" ~ "6 to 12 months",
      type == "year" ~ "1 to 3 years",
      type == "years" & strtoi(numb) < 4 ~ "1 to 3 years",
      type == "years" & strtoi(numb) < 7 ~"4 to 6 years",
      type == "years" ~"7+ years"
  ))

adoption_data <- adoption_data %>%
  separate(outcomeAge, c("numb", "type")) %>% 
  mutate(
    outcomeAgeBins = case_when(
      type == "day" ~ "Less than 6 months",
      type == "days" ~ "Less than 6 months",
      type == "week" ~ "Less than 6 months",
      type == "weeks" ~ "Less than 6 months",
      type == "month" ~ "Less than 6 months",
      type == "months" & strtoi(numb)< 6 ~ "Less than 6 months",
      type == "months" ~ "6 to 12 months",
      type == "year" ~ "1 to 3 years",
      type == "years" & strtoi(numb) < 4 ~ "1 to 3 years",
      type == "years" & strtoi(numb) < 7 ~"4 to 6 years",
      type == "years" ~"7+ years"
  ))

```

```{r crete duration bins}
adoption_data <- adoption_data %>%
  mutate(
    timeBetweenIntakeAndOutcomeBins = case_when(
      timeBetweenIntakeAndOutcome < 7 ~ "Less than 1 week",
      timeBetweenIntakeAndOutcome < 30 ~ "Less than 1 month",
      timeBetweenIntakeAndOutcome < 90  ~ "1 to 3 months",
      timeBetweenIntakeAndOutcome < 180 ~ "3 to 6 months",
      timeBetweenIntakeAndOutcome < 365 ~ "6 to 12 months",
      timeBetweenIntakeAndOutcome >= 365  ~ "1+ years"))
```

```{r crete gender}
adoption_data <- adoption_data %>%
  mutate(
    gender = case_when(
      grepl("female",intakeSex,ignore.case = TRUE)~ "Female",
      grepl("male",intakeSex,ignore.case = TRUE)~ "Male"))
```

```{r remove animals without a gender }
adoption_data <- subset(adoption_data, gender == "Male" | gender == "Female")
```


```{r remove missing values}
adoption_data[complete.cases(adoption_data),]
```

```{r distribution of animal types}
ggplot(data = adoption_data, mapping = aes(x = animalType)) +
  geom_bar()
```

Based on the very few number of non cat/dog animals I chose to remove non-cat/dogs from the dataset. 

```{r remove non-cat/dogs}
adoption_data <- subset(adoption_data, animalType == "Cat" | animalType == "Dog")
```


```{r distribution of dogs/cats by gender}
ggplot(data = adoption_data, mapping = aes(x = animalType)) +
  geom_bar()
```
Removing the other animals from the dataset gives us the distribution of dogs and cats seen above. We can see that there are about twice as many dogs in this dataset then there are cats.

```{r distribution of dogs/cats by gender}
ggplot(data = adoption_data, mapping = aes(x = animalType, fill=gender)) +geom_bar(position="stack")
```
Looking at the distribution of dogs and cats by gender we can see that there the genders seem to be distributed rather evenly across animal types


```{r intake count by month}
positions = c(1,2,3,4,5,6,7,8,9,10,11,12)
ggplot(data = adoption_data, mapping = aes(x = intakeMonth)) +  scale_x_discrete(limits = positions)+
  geom_bar()

```

```{r outcome count by month}
positions = c(1,2,3,4,5,6,7,8,9,10,11,12)
ggplot(data = adoption_data, mapping = aes(x = outcomeMonth)) +  scale_x_discrete(limits = positions)+
  geom_bar()
```

```{r distribution of intakeTypes}
ggplot(data = adoption_data, mapping = aes(x = intakeType)) +
  geom_bar()
```


```{r distribution of outcome types}
ggplot(data = adoption_data, mapping = aes(x = outcomeType)) + scale_x_discrete(labels=wrap_format(10))+
  geom_bar()
```

```{r distribution of intaketypes split by outcome type}
ggplot(data = adoption_data, mapping = aes(x=intakeType,fill=outcomeType))+geom_bar(position="dodge") +scale_x_discrete( labels=wrap_format(8))
```


```{r distribution of gender by outcome type}
ggplot(data = adoption_data, mapping = aes(x = gender)) + scale_x_discrete(labels=wrap_format(8))+
  geom_bar() + facet_wrap(~outcomeType)
```


```{r distribution of animal type by outcome type}
positions <- c("Adoption","Died","Euthanasia", "Return to Owner","Transfer")

ggplot(data = adoption_data, mapping = aes(x = outcomeType)) + scale_x_discrete(limits = positions, labels=wrap_format(8))+
  geom_bar() + facet_wrap(~animalType) 
```


```{r intake age distibution}
positions <- c("Less than 6 months","6 to 12 months", "1 to 3 years", "4 to 6 years", "7+ years")

ggplot(data = adoption_data, mapping = aes(x = intakeAgeBins)) + scale_x_discrete(limits = positions) +
  geom_bar()
```

```{r outcome age distribution}

positions <- c("Less than 6 months","6 to 12 months", "1 to 3 years", "4 to 6 years", "7+ years")

ggplot(data = adoption_data, mapping = aes(x = outcomeAgeBins)) + scale_x_discrete(limits = positions) +
  geom_bar()
```

```{r distibution of time between intake and outcome}
positions <- c("Less than 1 week","Less than 1 month", "1 to 3 months", "3 to 6 months", "6 to 12 months", "1+ years")
ggplot(data = adoption_data, mapping = aes(x = timeBetweenIntakeAndOutcomeBins)) + scale_x_discrete(limits = positions, labels=wrap_format(8)) +
  geom_bar()
```

```{r distibution of time between intake and outcome by animal type}
positions <- c("Less than 1 week","Less than 1 month", "1 to 3 months", "3 to 6 months", "6 to 12 months", "1+ years")
ggplot(data = adoption_data, mapping = aes(x = timeBetweenIntakeAndOutcomeBins)) + scale_x_discrete(limits = positions, labels=wrap_format(8)) +
  geom_bar() + facet_wrap(~animalType)
```

```{r distibution of time between intake and outcome by gender}
positions <- c("Less than 1 week","Less than 1 month", "1 to 3 months", "3 to 6 months", "6 to 12 months", "1+ years")
ggplot(data = adoption_data, mapping = aes(x = timeBetweenIntakeAndOutcomeBins)) + scale_x_discrete(limits = positions, labels=wrap_format(5)) +
  geom_bar() + facet_wrap(~gender)
```



```{r distibution of time between intake and outcome by outcomeType}
positions <- c("Less than 1 week","Less than 1 month", "1 to 3 months", "3 to 6 months", "6 to 12 months", "1+ years")


ggplot(data = adoption_data, mapping = aes(x=timeBetweenIntakeAndOutcomeBins,fill=outcomeType))+ scale_x_discrete(limits = positions, labels=wrap_format(8))+geom_bar(position="dodge") 
```


```{r}
ggplot(data = adoption_data, mapping = aes(x=timeBetweenIntakeAndOutcome))+geom_boxplot() 
ggplot(data = adoption_data, mapping = aes(x=timeBetweenIntakeAndOutcome))+geom_histogram() 
```





Is it possible to predict an animals outcome type based on the data that an adoption center can verify when intaking an animal?